---
title: "Vaccine simulations"
output: html_document
date: "2024-01-16"
---

```{r setup, include=FALSE, warning=FALSE}

library(tidyverse)
options(width = 100)

```

## Introduction

We are interested in doing an analysis to look at effectiveness of different vaccines (i.e. monovalent vs bivalent) and effects. We have estimates of population by age, when they were vaccinated and a linelist of positive cases with their vaccine history. Before we fit to data, we are going to simulate some fake data to make sure we know what we are doing.

## Simulation

### Basic Sim 

We are going to start from the basics and build our way up. We first start with a population and we vaccinate a group of them and compare the if the proportion of infections. Let's assume the vaccine has VE = p and the proportion of people infected in the unvaccinated group is x. Then the proportion of vaccined that's infected is x times (1-p). 

- population = 10,000
- proportion vaccinated = 50%
- VE = 80%
- FOI = 5%

```{r basic_params,echo=FALSE}
pop <- 1e4
vprop <- 0.5
ve <- 0.8
foi <- 0.05
```


```{r basic_sim, echo=FALSE}
basic_sim <- function(pop,vax_prop,ve,foi){
	dd <- data.frame(vax = c("unvax","vax")
		, size = pop*c(1-vax_prop, vax_prop)
		, prop = c(foi,foi*(1-ve))
	)
	return(dd)
}

basic_dat <- basic_sim(pop=pop,vax_prop=vprop,ve=ve,foi=foi)

print(basic_dat)

```

### Basic fit

We are now going to fit it now. 

```{r basic_fit}

basic_fit <- glm(prop~vax
	, weights = size
	, data = basic_dat
	, family = binomial()
)

print(summary(basic_fit))

```
We are going to use emmeans to get the effects.

```{r basic_emmeans}
library(emmeans)

basic_emm <- emmeans(basic_fit, ~vax, type = "response")

gg <- plot(basic_emm, type = "response")

print(gg)

```

Now we are going to try to recover VE using contrast. 

```{r basic_contrast}

basic_con <- contrast(basic_emm)

print(basic_con)

## Setting up the contrast

basic_con2 <- contrast(basic_emm
	, list(vax_vs_unvax = c(-1,1)
	)
)

print(basic_con2)

print(plot(basic_con2))
```

Recovering 0.8:

```{r calc_eff}
cc <- coef(basic_fit)
1- exp(cc[2])
1 - plogis(cc[1]+cc[2])/plogis(cc[1])
```

